### 2023-04-19 latex/paper.tex
Writin

\documentclass{siamonline220329}
%\documentclass{amsart}
\input{definitions}

%\numberwithin{equation}{section}

\title{Sensor Clusterization in D-optimal Designs for Bayesian Linear
  Inverse Problems}

\author{Yair Daon\thanks{Porter School of the Environment and Earth
    Sciences, Tel Aviv University, Tel Aviv, Israel
    (\email{yair.daon@gmail.com}).}}
%% \address{Porter School of the Environment and Earth
%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{}
%% \email{yair.daon@gmail.com}
% The 2010 edition of the Mathematics Subject Classification is the
% current definitive version.
%% \subjclass{Primary: 
%%   62F15, % Statistics - bayesian inference
%%   35R30, % PDE - Inverse problems
%%   Secondary:
%%   28C20,
%%   % Measure and integration - set functions and measures and integrals
%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
%%   % measure, etc.
%% }

\date{\today}
\begin{document}
\maketitle
\begin{abstract}
  We investigate the problem of sensor clusterization in D-optimal
  experimental design for infinite-dimensional Bayesian linear inverse
  problems. Many authors seem to believe that the fact that spatial
  correlations mitigateWe suggest an analytically tractable model for D-optimal
  designs. We argue that sensor clusterization in the case of iid
  observation noise is generic. We also prove that a model error term
  mitigates the clusterization phenomenon.

  %% As a part of the analysis we prove a Matrix Determinant
  %% Lemma in Hilbert spaces and a lemma for calculating
  %% $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
  %% operator-valued functions. We also show how to decompose $M = AA^t$
  %% with $A$ constrained to have unit norm columns.
\end{abstract}

\maketitle

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}

%% \section{Open Problems}

\section{Acknowledgements}
This study is a part of my PhD thesis. Most of it was written under
the instruction of Prof. Georg Stadler of the Courant Institute of
Mathematical Sciences. I would like to thank Prof. Stadler for his
great mentorship, attention to details and kind behavior.

This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE.  This work was also supported by The Raymond
and Beverly Sackler Post-Doctoral Scholarship.

\input{appendix}

\bibliographystyle{siamplain}
\bibliography{/home/yair/projects/library.bib}

\end{document}