### 2023-05-07 latex/paper.tex
Add abstract (raw)

\documentclass{siamonline220329}
%\documentclass{amsart}
%\documentclass{article}
\input{definitions}

%\numberwithin{equation}{section}

\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
  Inverse Problems}

\author{Yair Daon\thanks{Porter School of the Environment and Earth
    Sciences, Tel Aviv University, Tel Aviv, Israel
    (\email{yair.daon@gmail.com}).}}
%% \address{Porter School of the Environment and Earth
%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{}
%% \email{yair.daon@gmail.com}
% The 2010 edition of the Mathematics Subject Classification is the
% current definitive version.
%% \subjclass{Primary: 
%%   62F15, % Statistics - bayesian inference
%%   35R30, % PDE - Inverse problems
%%   Secondary:
%%   28C20,
%%   % Measure and integration - set functions and measures and integrals
%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
%%   % measure, etc.
%% }

\date{\today}
\begin{document}
\maketitle
\begin{abstract}
  \textbf{Setup} Taking measurements for inversion for physical
  processes requires expesive measurements.  \textbf{Problem}
  Unfortunately, when optimizing measurement locations, measurements
  get clustered.  \textbf{Why care} This a imortatn \textbf{What
    others do} Many researchers tried to avoind measuremnt
  clusterization, typically by modifying the optimization problem
  \textbf{What we do} We consider an analytically tractable model for
  Bayesian D-optimal design over Hilbert spaces. We prove that
  measurement clusterization poses no obstruction to D-optimality. We
  demonstrate that measuremnt clusterization is a generic property of
  D-optimal designs, and prove that measurement correlation mitigates
  the problem of measurement clusterization. We also give a full
  characterization of D-optimal designs, which, surprisingly, shows
  that D-optimal desings reduce uncertainty in the space of
  observations.  \textbf{Test on data} We test our conclusions via
  numeric implementation of our model
\end{abstract}

\maketitle

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}

%% \section{Open Problems}

\section{Acknowledgements}
This study is a part of my PhD thesis. Most of it was written under
the instruction of Prof. Georg Stadler of the Courant Institute of
Mathematical Sciences. I would like to thank Prof. Stadler for his
great mentorship, attention to details and kind behavior.

This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE.  This work was also supported by The Raymond
and Beverly Sackler Post-Doctoral Scholarship.

\input{appendix}

\bibliographystyle{siamplain}
%% %%\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/library.bib}

\end{document}