### 2023-05-10 latex/paper.tex
Improving

\documentclass{siamonline220329}
%\documentclass{amsart}
%\documentclass{article}
\input{definitions}

%\numberwithin{equation}{section}

\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
  Inverse Problems over Hilbert Spaces}

\author{Yair Daon\thanks{Porter School of the Environment and Earth
    Sciences, Tel Aviv University, Tel Aviv, Israel
    (\email{yair.daon@gmail.com}).}}
%% \address{Porter School of the Environment and Earth
%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{}
%% \email{yair.daon@gmail.com}
% The 2010 edition of the Mathematics Subject Classification is the
% current definitive version.
%% \subjclass{Primary: 
%%   62F15, % Statistics - bayesian inference
%%   35R30, % PDE - Inverse problems
%%   Secondary:
%%   28C20,
%%   % Measure and integration - set functions and measures and integrals
%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
%%   % measure, etc.
%% }

\date{\today}
\begin{document}
\maketitle
\begin{abstract}
  Inversion for parameteres of physical processes typically requires
  taking expensive measurements. Finding an optimal set of
  measurements is known as the optimal design problem. Surprisingly,
  measurement locations in optimal design are sometimes extremely
  clustered. Practically and intuitively, measurement clusterization
  is a huge problem. Researchers usually avoid measurement
  clusterization by somehow modifying the optimal design problem. We
  consider a certain flavor of the optimal design problem, based on
  the Bayesian D-optimality criterion. We suggest an analytically
  tractable model for Bayesian D-optimal designs in linear inverse
  problems over Hilbert spaces. We prove that measurement
  clusterization poses no obstruction to D-optimality. We demonstrate
  that measurement clusterization is a generic property of D-optimal
  designs, and prove that measurement correlation mitigates the
  clusterization of measurements. We also give a full characterization
  of D-optimal designs, which shows that D-optimal designs uniformly
  reduce uncertainty in the space of observations. Most importantly,
  we show how measurement clusterization can be avoided without
  changing the optimal design problem.
\end{abstract}

\maketitle

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}

%% \section{Open Problems}

\section{Acknowledgements}
This study is a part of my PhD thesis. Most of it was written under
the instruction of Prof. Georg Stadler of the Courant Institute of
Mathematical Sciences. I would like to thank Prof. Stadler for his
great mentorship, attention to details and kind behavior.

This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE.  This work was also supported by The Raymond
and Beverly Sackler Post-Doctoral Scholarship.

\input{appendix}

\bibliographystyle{siamplain}
%% %%\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/library.bib}

\end{document}