### 2023-05-10 latex/reply.tex
Moving stuff around and saving

\documentclass{amsart}
\input{definitions}
\numberwithin{equation}{section}
\begin{document}

\section{Referee 1}

\subsection{Overview}

This article discusses D-optimal experimental design for infinite
dimensional Bayesian linear inverse problems, and attempts to provide
a mathematical foundation to understanding the issue of clusterization
in such scenario, i.e., finding D-optimal designs.

As the author mentioned in the Acknowledgments Section, the paper is
paper is part of the Ph.D. dissertation, and in general provides a
good prospective to sensor placement for Bayesian linear inverse
problems formulated using infinite dimensional approach.  I regret,
however, to say that the paper requires a major revision before being
considered for publication.  Below, I provide some comments, that
hopefully will help improve the manuscript.  Also, provided are some
recent references, the author may find useful.

\subsection{General comments}

\begin{enumerate} 
  \item The author should clearly state why one should consider
  infinite-dimensional formulation of the observation operator? One
  could just merge the observation operator into the forward simulation
  operator, and follow standard formulations adopted by the works in the
  references below.
  \A{Thanks for the comment. I added clarification on the subject.}

  \item Moreover, when it comes to sensor placement, observations are
    almost surely pointwise in space/time. In this case, I don't see
    the benefit of consider infinite-dimensional formulation.
    \A{Indeed, observations are pointwise. I had to revert to a
    Hilbert space setting in order to carry out the analysis.}

  \item The contributions detailed in Subsection 1.2 (page 2) are very
    much appreciated, however they raise some concerns:
    \begin{description}
      \item [a] line 30: the author claims to present an analytically
        tractable model for understanding D-optimal designs. By
        reading the paper, I couldn't figure out how this reflects to
        any practical scenario. I'm not sure how this can be utilized
        for example in spatial sensor placement problems. An
        algorithmic description would be very useful here.  \A{I plan
          to pursue applications in following studies. Here I only
          aimed to understand sensor clusterization. In this respect
          you are correct --- there are no immediate/ apparent
          applications to the current study.}

      \item [b] The statement made on page 2, lines 40-41 (... apply
        equally to any of the scenarios...) is rather strong and
        unjustified.
        \A{Agreed. I removed said statement.}
    \end{description}

    \item On page 8, Equation 4.2, which is crucial in this study,
      assumes that the forward operator is invertible. This is clearly
      impossible in practical Bayesian inverse problems, otherwise, it
      reduces to a standard regression problem. This is understandable
      for clarification, however, the following argument on Page 9
      that extends this assumption is not well-justified.
      \A{Thanks for this comment, I was able to remove the invertibility
      assumption.}

    \item In Equation 4.3, not sure what $^{-1}$ stands
      for. Moreover, an inverse of the forward operator is not
      well-defined even if $\fwd\Gamma_{prior}\fwd$ is invertible!
      This is very limiting, and can't be used in practical inverse
      problems. This should be clearly discussed for the reader to
      appreciate the valuable effort made.
      \A{I thought $^{-*}$ is common notation to "inverse
        adjoint". Clearly I was mistaken. I added clarification.}

    \item A numerical example with the setup studied in the paper,
      should help clarify several confusions that a reader can fall
      for easily.
      \A{TODO}

    \item The title should reflect the fact that the study is
      restricted to linear inverse problems. Also, this should be
      clarified in the abstract.
      \A{Done.}

    \item In addition to the fact that observations errors are
      uncorrelated with fixed variance $\sigma^2$.

      \A{Added emphasis for linear inverse problems in the title and
      abstract, thanks!! However, observation errors in the model
      presented are not always uncorrelated. I address this in the
      model correlation term.}
\end{enumerate}

\subsection{Other Comments}

\begin{enumerate}
\item In the abstract (line 20, page 1), and on page 5, lines 30-40, the
  author mentions that the effect of observation-correlations on sensor
  clusterization will be considered later, but I failed to find any
  relevant discussion to that issue which I find interesting.

  \A{This happens in section blabla. I now refer to this in the
    manuscript.}

  \item Even if the model error covariance $\Gamma_{model}$ is known,
  this doesn't amount for observation correlations, .e.g,
  single-instrument-observations, because the observation operator
  $\mathcal{O}$ is non invertible. This issue should addressed and
  clarified clearly.

  \item I found it difficult to trace the referenced Lemmas,
    Corollaries, etc., mainly because bad referencing in the appendix,
    and the links do not works properly.  For example Corollary 2
    should be Corollary B.x, etc.

    \A{Thanks, I made the lemmase numbering consistent.}

  \item The trace operator should be followed with some sort of brackets to
    make it easier to follow and validated the formulae

    \A{Added, thanks!}

  \item Sensor clusterization in OED is central to this study. Figure
    2 is used to illustrate the phenomena, however, it is
    underrepresented, and is hard to decipher. Specifically, Section
    4.2 (Page 10, lines 36-39), and caption of Figure 2 should be
    rewritten with clear explanation of the axes, and the
    plot-generation process.

    \A{Fuck you, you cunt!}

    \item One page 10, lines 30-35, the argument doesn't seem to be
      quite accurate. It's not always true that one associates a
      higher weight/value to ``better'' observation. If one defines
      ``better'' as lower uncertainty, then, variable observation
      variances must be considered. On the other hand, if being
      ``better'' corresponds to a more accurate realization of the
      observation (true snapshot of reality), then the whole statement
      would be wrong, because optimal design relies only on the
      uncertainty levels in linear Bayesian settings, as manifested by
      the posterior covariance matrix.

    \item The last statement of the proof of Lemma A.1 (see page 12,
      line 24), stating that ``there is nothing special about $t=0$,
      and the relation holds for all t''. This is rather odd. In
      fact, It is assumed that $X(t=0)$ explicitly (see line 32 on
      page 11). A clarification is required here.

      \A{This is actually how the original proof of Lax
      ends. Regardless, I added clarification.}
\end{enumerate}

\subsection{Minor Comments}

\begin{enumerate}

  \item Page 1, line 22; ... ``calculating derivative of log det of
    oeprators''. Please clarify for the reader, which operators you're
    referring to.

  \item Page 2, line 11; wrong reference to Figure 1.3?

  \item Page 6; in equation 2.7, don't you need an expectation over
    all priors to achieve this result? Please clarify.

  \item Page 6, line 13; the optimization problem is solved for
    $\mathcal{O}$, not $\Psi$ ?

  \item Page 6, lines 19; the statement made about relaxing the
    constraints on the observation operator is unclear.

  \item Page 6, line 24, I believe the reference to (3.2) should be
    replaced with (3.1).

  \item Page 6, line 26, proof is provided in which Appendix?
    Please state clearly.

  \item page 6, In the proof of Lemma 3.1, the author utilized the
    operator $T$, which is only introduced in the appendix (Proof of
    Lemma A.1). Maybe state that clearly here. Also, in the proof,
    from step 4 to 5, the first and third terms vanished; why?

  \item Page 7; Proposition 1 should be replaced with something like
    Proposition 3.2 or something similar to match previous
    referencing (e.g., Lemma 3.1).

  \item Page 7, line 27 (and also line 31); the reference to Proposition 2
    seems to be incorrect. Also, state clearly which appendix the proof is
    delegated to.

  \item Page 10, Corollary 1 should be renamed (see the comment on
    Proposition 1 above). Same applies to Corollary 3 in the
    appendix. Better also to clarify which appendix. Finally, the
    author utilizes $\widehat{\Gamma_{post}}$ int the corollary,
    which is defined only in the appendix, under the proof.

  \item Page 11, in the proof of Lemma A.1 (lines 39-43), the partial
    derivative vanished from step 2 to step 3; please explain!

\end{enumerate}
\subsection{References}

\begin{itemize}
\item Koval, Karina, Alen Alexanderian, and Georg Stadler. "Optimal
  experimental design under irreducible uncertainty for linear inverse
  problems governed by PDEs." Inverse Problems (2020).

\item Attia, Ahmed, and Emil Constantinescu. "Optimal Experimental
  Design for Inverse Problems in the Presence of Observation
  Correlations." arXiv preprint arXiv:2007.14476 (2020).
\end{itemize}

Letter reference: DSRR01

\section{Referee 2}

\subsection{Summary}

In this article, the author considers what causes sensor clustering in
optimal experimental design of Bayesian inverse problems. They show
how D-optimal designs may lead to sensor clustering in the case of iid
measurement noise. They also show that clustering does not occur for
spatially correlated measurement error. The paper is somewhat
difficult to follow due to writing and organizations corrections that
should be made, but does present original work. The article is
mathematically rigorous and significant in that the work is a
necessary step toward developing a deeper understanding of sensor
clustering. I suggest a major review and resubmission after reviewing
the comments below.

\subsection{General Comments}

1. It should be stated in the Abstract whether the work in this paper
pertains to only linear inverse problems, or if it is applicable in
general to non-linear inverse problems as well.

\A{Added clarification, this paper only pertains to linear inverse
  problems.}

2. It is unclear from the contribution section of the article what
exactly the primary contribution is. This should be stated very
explicitly and empasized clearly as the contribution of the article.

3. In some equations (as seen below in the line comments) operators
are not clearly defined or described before they are given in a
proposition or theorem. Please clearly define all operators both with
mathematical definitions and descriptions (when applicable) before
using them.

4. Overall, I find the article difficult to follow. I think this can
be improved by properly defining all operators and including section
overviews that give a description of the purpose of each section, and
a brief summary of how this is being accomplished.

5. On page 10, figure 2 and section 4.2 deliver the main conclusion of
the article. These areas are deficient with only brief descriptions,
and should be much more thoroughly explained. In particular page 10,
lines 38-39 "If the precision in these is lower than the precision for
the next eigenvector, a repeated measurement is optimal" needs to be
elaborated on. How do you come to this conclusion? Additionally, the
wording of the statement can lead to confusion. From figure 2, it
looks like all measurements grant less precision than that of the
fourth eigenvector, so I first interpret this as all measurements
should be repeated. I believe these issues can be fixed by significant
elaboration in section 4.2.

\subsection{Line Comments}

Page 1, Line 39: "In a frequentist settings" should be changed to "in
a frequentist setting" or "in frequentist settings. \A{Changed.}

Page 2, Line 36: Add a comma: "In the process, we generalize..." \A{Added.}

Page 3, Line 37: "with homogeneous Dirichlet boundary condition"
should be changed to "with a homogeneous Dirichlet boundary
condiiton". \A{Added.}

Page 4, Line 14: In equation (2.1), $\Sigma$ has not been
defined. Please formally define this before giving the equation.

Page 4, Line 21: The prior distribution $\mathcal{N}(0,\Gamma_{pr})$
assumes the prior mean is 0. This is worthy of some comment. Why must
the prior mean be 0? \A{Indeed, the prior does not need to equal
  zero. I changed this in the manuscript.}

Page 5, Line 15: Add a comma: "Considering the sum of the error terms,
it is easy..." \A{Added.}

Page 6, Line 7: Please describe what $\Psi$ is in the theorem. It is
defined mathematically but not described. Only much later is it stated
that this is the objective function. \A{Added a definition for a
D-optimal design, I think this explains what $\Psi$ is and why we care
about it.}

Page 6, Line 19-20: "In section 3.2 we relax the constrains on
$\mathcal{O}$ so that the optimization problem in (3.1) and find
gradients for the new constraints." This sentence is unclear and
should be reworded.

Page 7, Line 24, 27, 31: The proposition is labeled as Proposition 1,
but it is refered to as Proposition 2, and relabeled Proposition 2 in
the appendix. \A{Fixed, thanks!}

Page 9, Line 6: Please be more exact when refering to "the big
parentheses" to avoid any chance of confusion. Give the explicit
formula here as you do for another term earlier in the
sentence. \A{Done, thanks!}

Page 18, Line 6: This proposition was previously labeled Proposition
1. \A{Fixed, thanks!}
\end{document}