### 2023-05-10 latex/tikz_clusterization.tex
Clusterization illustration with my code

\begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
  \begin{axis}
    %% title={Posterior Pointwise Standard Deviations and D-Optimal
    %% Measurement Locations}
    [xmin = 0, xmax = 3.14, xlabel = {$x \in \Omega$}, ylabel =
      posterior pointwise std, ymin = 0,
    %compat = 1.3,
    % ymax   = 130,
    % ytick = \empty,
    legend cell align=left,
    % legend style={at={(0.45,0.2)}}
    legend pos= outer north east 
    ]
    % \draw[black!30!white, thin] (50,0) -- (50,130);
    % 
    %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
    %% \addlegendentry{1 measurements};

    %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
    %% \addlegendentry{2 measurements};

    %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
    %% \addlegendentry{3 measurements};

    \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
    \addlegendentry{4 measurements};

    \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
    \addlegendentry{5 measurements};

    \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
    \addlegendentry{6 measurements};

    %% \addplot [black,  only marks, mark=x, mark size=1.5] 
    %% table{locs-heat-sens1-var1.txt}; 
    %% \addplot [blue,   only marks, mark=x, mark size=1.5]
    %% table{locs-heat-sens2-var1.txt}; 
    %% \addplot [red,    only marks, mark=x, mark size=1.5]
    %% table{locs-heat-sens3-var1.txt};
    \addplot [green,  only marks, mark=*, mark size=1.5] 
    table{locs-heat-sens4-var1.txt}; 
    \addplot [purple, only marks, mark=*, mark size=1.5] 
    table{locs-heat-sens5-var1.txt}; 
    \addplot [cyan,   only marks, mark=*, mark size=1.5] 
    table{locs-heat-sens6-var1.txt};

  \end{axis}
\end{tikzpicture}