### 2023-05-14 latex/paper.tex
Done?

\documentclass{siamonline220329}
%\documentclass{amsart}
%\documentclass{article}
\input{definitions}

%\numberwithin{equation}{section}

\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
  Inverse Problems over Hilbert Spaces}

%% \author{Yair Daon\thanks{Porter School of the Environment and Earth
%%     Sciences, Tel Aviv University, Tel Aviv, Israel
%%     (\email{yair.daon@gmail.com}).}}
\author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
    University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}

%% \address{Porter School of the Environment and Earth
%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{}
%% \email{yair.daon@gmail.com}
% The 2010 edition of the Mathematics Subject Classification is the
% current definitive version.
%% \subjclass{Primary: 
%%   62F15, % Statistics - bayesian inference
%%   35R30, % PDE - Inverse problems
%%   Secondary:
%%   28C20,
%%   % Measure and integration - set functions and measures and integrals
%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
%%   % measure, etc.
%% }

\date{\today}
\begin{document}
\maketitle
\begin{abstract}
  Inversion for parameteres of physical processes typically requires
  taking expensive measurements; finding an optimal set of
  measurements is known as the optimal design problem. Surprisingly,
  measurement locations in optimal designs are sometimes extremely
  clustered. Since measurement clusterization is nonintuitive,
  researchers often avoid it by modifying the optimal design
  problem. We consider a certain flavor of the optimal design problem,
  based on the Bayesian D-optimality criterion, and suggest an
  analytically tractable model for Bayesian D-optimal designs in
  linear inverse problems over Hilbert spaces. We prove that
  measurement clusterization poses no obstruction to D-optimality. We
  demonstrate that measurement clusterization is a generic property of
  D-optimal designs, and prove that correlated noise between
  measurements mitigates clusterization. We also give a full
  characterization of D-optimal designs, which shows that D-optimal
  designs uniformly reduce uncertainty in the space of
  observations. Finally, we show how measurement clusterization can be
  avoided without changing the optimal design problem.
\end{abstract}

\maketitle

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}

%% \section{Open Problems}

\section{Acknowledgements}
This study is a result of research I started during my PhD studies
under the instruction of Prof. Georg Stadler of the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kind behavior. This work was
supported by The Raymond and Beverly Sackler Post-Doctoral
Scholarship.

This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE.

%% \input{appendix}

\bibliographystyle{siamplain}
%% %%\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/library.bib}

\end{document}