### 2023-05-19 latex/how.tex
add explanation for why clusterization

%\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
\section{Substituting an optimal clustered design with an equally optimal non-clustered design}\label{section:how}
%
The necessary conditions for D-optimality of Theorem \ref{thm:char}
leave some freedom in choosing $\obs$, which might allow us to avoid a
clustered design. As an example, consider an inverse problem for which
the first two eigenvalues of the pushforward prior $\fwd \prcov\fwd^*$
are $\lambda_1 = 10, \lambda_2 = 5$, with eigenvectors
$\{\ev_i\}_{i=1}^\infty$ and $\sigma^2=1$. Consider the following
$\obs, \pobs$ (see also Figure \ref{fig:clusterization}):
\begin{equation}\label{eq:two designs}
  \obs =
  \begin{pmatrix}
    \sqrt{\frac{11}{40}}\ev_1 - \sqrt{\frac{29}{40}}\ev_2 \\
    \sqrt{\frac{11}{40}}\ev_1 + \sqrt{\frac{29}{40}}\ev_2 \\
    \ev_1 \\
    \ev_2 \\
    \ev_1 \\    
  \end{pmatrix},
  \pobs =
  \begin{pmatrix}
    \sqrt{\frac{3}{8}}\ev_1 - \sqrt{\frac{5}{8}}\ev_2 \\
    \sqrt{\frac{3}{8}}\ev_1 + \sqrt{\frac{5}{8}}\ev_2 \\
    \sqrt{\frac{2}{5}}\ev_1 - \sqrt{\frac{3}{5}}\ev_2 \\
    \sqrt{\frac{2}{5}}\ev_1 + \sqrt{\frac{3}{5}}\ev_2 \\
    \ev_1 \\    
  \end{pmatrix}
\end{equation}
\noindent It is easy to verify that $\obs^*\obs = \pobs^*\pobs$
and that both $\obs$ and $\pobs$ are D-optimal. However, $\obs$
is a clustered design ($\meas_3 = \meas_5 = \ev_1$), while $\pobs$ is
not.

We may now answer \textbf{Question \ref{q:replace}}: We showed that it
is possible to substitute a D-optimal clustered design (e.g. $\obs$)
for a D-optimal non-clustered design (e.g. $\pobs$). Presenting a
general algorithm for replacing a clustered D-optimal design with a
non-clustered D-optimal design is out of the scope of the current
study. We can, however, provide some intuition, utilizing $\obs,
\pobs$ as a test case. Let $\obs_{12}$ the design composed of the
first and second rows of $\obs$ (with a similar definition for
$\pobs_{12}$), and note that, up to permutations and sign changes,
$\obs_{12}$ uniquely determines $\obs$ as a clustered design. We find
$\pobs_{12}$ by requiring that (a) measurements in $\pobs_{12}$ have
unit norm, (b) $\pobs_{12}^* \pobs_{12}$ has the same eigenvectors as
$\obs^*\obs$, (c) $\pobs^*\pobs \approx \obs^*\obs$, and (d)
$\pobs^*\pobs \neq \obs^*\obs$. We then consider a ``new'' inverse
problem, for which the pushforward prior covariance is the covariance
of the posterior $\mu_{\textup{post}}^{\data, \pobs_{12}}$, namely

$$
\left(\left(\fwd \prcov\fwd^*\right)^{-1} + \sigma^{-2}
\pobs_{12}^*\pobs_{12}\right)^{-1}.
$$
%
Finally, we complete $\pobs$ to a D-optimal design with $m=5$
measurements via the construction outlined in Lemma
\ref{lemma:free}.

Interestingly, part (c) in the process detailed above implies that
there are actually infinitely many non-clustered D-optimal designs. It
is not at all clear why the numerical implementation of Lemma
\ref{lemma:free} results in a clustered design. An answer to this
question may further help shed light on the measurement clusterization
phenomenon.

\pgfplotstableread{
  Label    prior  $o_1o_2$  $o_3$   $o_4$   $o_5$    topper
  1         0.1      0.55       1       0       1      0.001
  2         0.2      1.45       0       1       0      0.001
  3         3.5      0          0       0       0      0.001
}\clusterization

\pgfplotstableread{
  Label    prior  $o_1o_2$  $o_3o_4$   $o_5$     topper
  1         0.1    0.75     0.8        1         0.001
  2         0.2    1.25     1.2        0         0.001
  3         3.5    0        0          0         0.001
}\noclusterization

\begin{figure}
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\clusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=red!60]    table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=black!60]  table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=orange!60] table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
      %% \addplot [fill=cyan!60]   table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
      %% \addplot [fill=purple!60] table [y=$o_7$, meta=Label, x expr=\coordindex] {\clusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      %% \addlegendentry{$o_6$}
      %% \addlegendentry{$o_7$}   
    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\noclusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=red!60]    table [y=$o_3o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=black!60]  table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1o_2$}
      \addlegendentry{$o_3o_4$}
      \addlegendentry{$o_5$}
      %% \addlegendentry{$o_4$}
      %% \addlegendentry{$o_5$}
      %% \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Clusterization (left, $\obs$ in \eqref{eq:two designs})
    and non-clusterization (right, $\pobs$ in \eqref{eq:two designs})
    in D-optimal designs. Posterior precision per eigenvector is
    plotted for $\obs$ and $\pobs$ from \eqref{eq:two
      designs}. Both designs are identical for all practical matters
    --- their posteriors are equal. In the left panel, a D-optimal
    design with clusterization is illustrated: $\meas_3 = \meas_5$. In
    the right panel, a D-optimal design without clusterization is
    illustrated. The contribution to the posterior precision due to
    $\meas_1$ and $\meas_2$ is denoted $\meas_1\meas_2$. These
    contributions are plotted together because neither $\meas_1$ nor
    $\meas_2$ are not in the direction of any eigenvector. Similarly
    for $\meas_3\meas_4$.}
  \label{fig:clusterization}
\end{figure}