### 2023-05-21 latex/why.tex

\section{An answer for Question \ref{q:why}: The cause of measurement clusterization}\label{section:why}
\begin{figure}[H]
    \centering
    \includegraphics[height=0.5\textwidth]{eigenvectors.eps}
    \caption{D-optimal measurement locations ($m=5$ measurements) and
      weighted eigenvectors for inversion of the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from a zero of the
      first and second eigenvectors. Allocating $m=5$ measurements
      into two locations results in clusterization, according to the
      pigeonhole principle.}
  \label{fig:why}
\end{figure}

According to Theorem \ref{thm:char}, D-optimal designs aim to capture
a small subset of eigenvectors of the prior covariance, specifically
the $k$ eigenvectors with the highest prior variance. Our model
naturally achieves this objective by not measuring eigenvectors $k+1$
and above, as proven in Theorem \ref{thm:char}. Translating this
understanding to spatial problems, we anticipate that a D-optimal
design would favor measurement locations where eigenvectors $k+1$ and
above are either close to zero in value or possess small eigenvalues
in the prior spectrum, for some $k > 0$. To illustrate this
preference, Figure \ref{fig:why} depicts the scenario using the 1D
heat equation with homogeneous Dirichlet boundary
conditions\footnote{Additional details provided in the supplementary
material.}. The plot showcases four eigenvectors, scaled according to
their prior standard deviations. Since eigenvectors beyond the fourth
have insignificant prior eigenvalues, we exclude them from
consideration. Notably, we observe that measurements are clustered
near the zeros of the third and fourth eigenvectors, so we conclude
that $k=2$. Clusterization arises because there are only two locations
where the third and fourth eigenvectors approach zero, while the first
and second eigenvectors exhibit significantly non-zero
values. Consequently, when we allocate four measurements to these two
locations, they naturally cluster together, aligning with the
pigeonhole principle.