### 2023-05-21 arxiv/ms.tex
Ready for arxiv

%\documentclass{siamonline220329}
\documentclass{amsart}
\input{definitions}

%\numberwithin{equation}{section}

\title[Measurement Clusterization in D-optimal Designs]{Measurement
  Clusterization in D-optimal Designs for Bayesian Linear Inverse
  Problems over Hilbert Spaces}

\author{Yair Daon}

%% \thanks{Porter School of the Environment and Earth
%%     Sciences, Tel Aviv University, Tel Aviv, Israel
%%     (\email{yair.daon@gmail.com}).}
%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}

\address{Porter School of the Environment and Earth Sciences, Tel Aviv
  University\\ Tel Aviv, Israel}
\address{Sackler Faculty of Medicine,
  Sciences, Tel Aviv University\\ Tel Aviv, Israel}

%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
%% %% \curraddr{}
\email{yair.daon@gmail.com}
%% The 2010 edition of the Mathematics Subject Classification is the
%% current definitive version.

\subjclass{Primary: 
  62F15, % Statistics - bayesian inference
  35R30, % PDE - Inverse problems
  Secondary:
  28C20.
  % Measure and integration - set functions and measures and integrals
         % in infinite-dimensional spaces (Wiener measure, Gaussian
  % measure, etc.
}

\date{\today}
\begin{document}

\begin{abstract}
  Inversion for parameters of physical processes typically requires
  taking expensive measurements; finding an optimal set of
  measurements is known as the optimal design problem. Surprisingly,
  measurement locations in optimal designs are sometimes extremely
  clustered. Since measurement clusterization is nonintuitive,
  researchers often avoid it by modifying the optimal design
  problem. We consider a certain flavor of the optimal design problem,
  based on the Bayesian D-optimality criterion, and suggest an
  analytically tractable model for Bayesian D-optimal designs in
  linear inverse problems over Hilbert spaces. We demonstrate that
  measurement clusterization is a generic property of D-optimal
  designs, and prove that correlated noise between measurements
  mitigates clusterization. We also give a full characterization of
  D-optimal designs, which shows that D-optimal designs uniformly
  reduce uncertainty in a select subset of the prior covariance
  eigenvectors. Finally, we show how measurement clusterization is a
  simple consequence of the characterization mentioned above and the
  pigeonhole principle.
\end{abstract}
\maketitle

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

\section{Acknowledgements}
This study is a result of research I started during my PhD studies
under the instruction of Prof. Georg Stadler at the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kindness. I would also like to
thank Christian Remling, who helped me find a proof for Lemma
\ref{lemma:free} in
\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.

This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE. This work was also supported by The Raymond
and Beverly Sackler Post-Doctoral Scholarship.

%% \input{appendix}

%\bibliographystyle{siamplain}
\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}