### 2023-08-10 latex/prelims.tex

\section{Preliminaries and Notation}\label{section:prelim}

In this section we present the setup that will be used throughout this
article. The theoretical foundations for inverse problems over
function spaces can be found elsewhere \cite{Stuart10} and will not be
reviewed here.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
and o are for "parameter" and "observation", respectively), and let
\(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
operator \(\fwd\) is assumed linear and strongly smoothing (the heat
operator of the 1D heat equation introduced in Section
\ref{section:intro} is a prime example). Take a Gaussian prior
\(\param \sim \pr = \normal(\prmean ,\prcov)\) with some appropriate
covariance operator \(\prcov\) on \(\hilp\) \cite{Stuart10}. Note that
\(\fwd \prcov \fwd^*\) is the prior covariance in \(\hilo\)
\cite{Stuart10}, and as such, \(\fwd \prcov \fwd^*\) is assumed
invertible --- an assumption which will be used below (if \(\fwd\) has
a nontrivial kernel we utilize Occam's Razor and ignore said kernel
altogether). Measurements are taken via a \emph{measurement operator}
\(\obs\). It is common for the measurement and forward operators to be
merged \(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but
the analysis carried out in the following sections requires that
\(\fwd\) and \(\obs\) are explicitly separated as in
\cite{attia2022stochastic, cvetkovic2023choosing}.  \(\obs \in (
\hilo^* )^m\), where \(m\) is the number of measurements
taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
\(\obs\) are called \emph{measurements}:

\begin{equation*}%\label{eq:O}
  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
\end{equation*}

Data is acquired via noisy observations, and we consider two types of
error terms: Spatially correlated model error \(\eps' \sim
\normal(0,\modcov)\) with \(\modcov\) a covariance operator. Observation
error is denoted \(\eps \sim \normal(0, \sigma^2 I_m)\), with \(I_m \in
\mathbb{R}^{m \times m}\) the identity. Both error terms and the prior
are assumed independent of each other. Thus, data is acquired via
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
\end{align}

It is easy to verify that \(\obs \eps' + \eps \in \R^m\) is a centered
Gaussian random vector with covariance matrix

\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps) (\obs \eps' +
      \eps)^t ]
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I_m , 
  \end{split}
\end{align}
where
\begin{align}\label{eq:modcov explained}
  \begin{split}
    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j 
    %
    %
    %
    = \meas_i (\modcov \meas_j).% \text{ (by \eqref{eq:obs*})}.
  \end{split}
\end{align}
Taking \(\modcov = 0\) is a common practice
\cite{tarantola2005,Kaipio2005,Vogel02} and then \(\Sigma =
\sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).

Finally, it is useful to record that the posterior measure \(\post\)
is Gaussian in this setting and its covariance operator does not
depend on data \(\data\) \cite{Stuart10}:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
divergence between posterior \(\post\) and prior measures \(\pr\). It is
useful to first recall the definition of KL divergence \cite{CoverThomas91}:
\begin{equation*}
  D_{KL}(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
\end{equation*}

The study of D-optimal designs for Bayesian linear inverse problems in
infinite dimensions was pioneered by \cite{AlexanderianGloorGhattas14,
  alexanderian2018efficient}. The main result we will make use of is
summarized (in our notation) below:

\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let \(\pr = \normal(\prmean,\prcov)\) be a Gaussian prior on \(\hilp\)
  and let \(\post = \normal(\postmean,\postcov)\) the posterior measure
  on \(\hilp\) for the Bayesian linear inverse problem \(\data = \obs
  \fwd\param + \obs \eps' + \eps\) discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}

Note that in \cite{AlexanderianGloorGhattas14,
  alexanderian2018efficient}, results are stated for \(\Sigma=I\)
(implied by \(\modcov = 0,\sigma^2=1\)), but these results also hold
for more general covariance matrices
\cite[p. 681]{AlexanderianGloorGhattas14}.

%% It is important to note that since \(\obs\) is finite-rank,
%% \(\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}\) is
%% trace-class.
\begin{definition}\label{def:d optimality}
  We say \(\obs^{\star}\) is \emph{D-optimal} if \(\obs^{\star} =
  \argmax_{\obs} \tar(\obs)\), where entries of \(\obs \in (\hilo^*)^m\)
  are constrained to some allowed set of measurements in \(\hilo^*\).
\end{definition}

Intuition for Theorem \ref{thm:d optimality} can be gained by
considering a Bayesian linear model in finite dimensions, with
Gaussian prior and Gaussian noise. Then, a D-optimal design minimizes
the determinant of the posterior covariance matrix, and this turns out
to be a regularized version of the frequentist D-optimality criterion
\cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition
\ref{def:d optimality} carry a similar intuition:
\begin{align*}
  \begin{split}
    \tar(\obs) &= \frac12 \log \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \text{ by \eqref{eq:objective}}\\
    &= \frac12 \log \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
    &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}}.
    %%  &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
  \end{split}
\end{align*}
We think of \(\prcov\) as constant, so a D-optimal design minimizes a
quantity analogous to the posterior covariance determinant, similarly
to the finite-dimensional case.

%% \subsection{Notation Summary}\label{subsec:notation}
%%   Our notation is summarized below. We let:
%%   \begin{itemize}
%%   \item \(\hilp, \hilo\) Hilbert spaces.
%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
%%     \in \mathbb{N}\) is the number of measurements taken.
%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
%%   \item \(\modcov\) model error covariance operator.
%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
%%   \item A D-optimality design criterion
%%     \cite{AlexanderianGloorGhattas14}:
%%     \begin{align*}
%%       \begin{split}
%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
%%         % 
%%         % 
%%         % 
%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
%%         \fwd \prcov^{1/2}).
%%       \end{split}
%%     \end{align*}
%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
%%   %%   in decreasing order of magnitude.
%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
%%   \end{itemize}

%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
%% \begin{align}\label{eq:optimization}
%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
%%   = \argmax_{\obs} \frac12 \log \det 
%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
%% \end{align}
%% where \(\obs\) is constrained to some allowed set of observations. We
%% call this problem ``simultaneous optimization'', since all
%% observations are decided on simulatneously.

%% For computational reasons, one may prefer to find the best
%% observations in a sequential manner. Denote
%% \begin{equation}\label{eq:def obs_k}
%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
%% \end{equation}
%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
%% found, where \(m\) is the number of available observations. %% It is
%% %% important to notice that this scheme does not require actually
%% %% observing data --- in \eqref{eq:objective} data is averaged out.

%% The analysis in this paper is conducted for the general simultaneous
%% optimization case. The sequential optimization case is dealt with in
%% section \ref{subsec:clusterization sequential}. It is important to
%% note, however that all conclusions we arrive at for the simultaneous
%% case easily specialize to the sequential case by considering the
%% posterior as the next sequential step's prior.