### 2023-08-29 latex/m3as.tex

\documentclass{ws-m3as}
\input{definitions}
\usepackage[super]{cite}
\usepackage{url}

\begin{document}

\markboth{Yair Daon}{Measurement Clusterization in D-optimal Designs}

%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Measurement Clusterization in D-optimal Designs for Bayesian
  Linear Inverse Problems over Hilbert Spaces}

\author{Yair Daon}

\address{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel\\
yair.daon@gmail.com}

\maketitle

\begin{history}
\received{(Day Month Year)}
\revised{(Day Month Year)}
%\accepted{(Day Month Year)}
\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
  \input{abstract}
\end{abstract}

\keywords{Optimal Design; Clusterization; Inverse Problem; Bayesian.}

\ccode{AMS Subject Classification: 62F15, 35R30, 28C20}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

\section{Acknowledgements}
\input{ack}

\input{funding}

\bibliographystyle{ws-m3as}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}