### 2023-10-01 latex/SIAM/siam.tex
Submit to Bayesian Analysis

\documentclass{siamart220329}
\input{definitions}

% Sets running headers as well as PDF title and authors
\headers{Measurement Clusterization in D-optimal Designs}{Yair Daon}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.

\title{Measurement Clusterization in D-optimal Designs for Bayesian
  Linear Inverse Problems over Hilbert Spaces
  \thanks{Submitted to the editors Aug. 29\textsuperscript{th} 2023}
  %% \funding{\input{funding}}
}

\author{Yair Daon \thanks{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
      Israel (\email{yair.daon@gmail.com}).}}

\begin{document}
\maketitle

\begin{abstract}
  \input{abstract}
\end{abstract}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

\section{Acknowledgements}
\input{ack}

%% \input{appendix}
\bibliographystyle{siamplain}
%\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}