### 2023-10-01 latex/ba.tex

% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
\input{definitions}
%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}

\startlocaldefs
% ** Local definitions **
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

%\title{\support{}}
\runtitle{Measurement Clusterization in D-optimal Designs}
%\author{Yair Daon}
%\numberwithin{equation}{section}

\begin{aug}
\author{\fnms{Yair} \snm{Daon}\thanksref{addr1}\ead[label=e1]{yair.daon@gmail.com}}
\runauthor{Yair Daon}

\address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
    \printead{e1} % print email address of "e1"
}

\end{aug}

\begin{abstract}
  \input{abstract}
\end{abstract}

%% ** Keywords **
\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd{35R30}
\kwd[; secondary ]{28C20}
\end{keyword}

\end{frontmatter}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}

\bibliographystyle{ba}
\bibliography{/home/yair/projects/bibtex.bib}

\begin{acks}[Acknowledgments]
\input{ack}

\input{funding}
\end{acks}

\end{document}