### 2023-10-01 latex/ms.tex

\documentclass{amsart}
\input{definitions}

\title[Measurement Clusterization in D-optimal Designs]{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

\author{Yair Daon}
\address{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel}

%\email{yair.daon@gmail.com}

%\numberwithin{equation}{section}

%% The 2010 edition of the Mathematics Subject Classification is the
%% current definitive version.
%% \subjclass{Primary: 
%%   62F15, % Statistics - bayesian inference
%%   35R30, % PDE - Inverse problems
%%   Secondary:
%%   28C20.
%%   % Measure and integration - set functions and measures and integrals
%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
%%   % measure, etc.
%% }

\begin{document}
\maketitle

\begin{abstract}
  \input{abstract}
\end{abstract}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}

\section{Acknowledgements}
\input{ack}
\input{funding}

\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}

%% \input{how}
%% \section{Open Problems}