### 2024-07-21 latex/ack.tex
before diff

This study is a result of research I started during my PhD studies
under the instruction of Prof.~Georg Stadler at the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kindness. I would also like to
thank Christian Remling, who helped me find a proof for Lemma
\ref{lemma:free} in
\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
Last, but certainly not least, I would like to thank the three
referees, associate editor and editor in chief for providing detailed
and insightful reviewes that have made this manuscript a whole lot
better.
%Having worked on this manuscript alone for several years, they have
%almost become my coautoh