### 2024-07-31 latex/refs.tex

\documentclass{amsart}
\input{definitions}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}

\newcommand{\paste}[1]{\vskip 0.05in \textcolor{red}{#1}\vskip 0.05in}
\newcommand{\answer}[1]{\vskip 0.05in \textcolor{blue}{#1} \vskip 0.05in}

\begin{document}

I appreciate the many thoughtful comments by all of you. I believe
these comments helped me considerably improve the exposition of my
paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit. Please see below my answers to the
comments made by the referees, associate editor and editor in chief:
\answer{My answers are marked in blue.} \paste{In some places I
  copied text from the revised manuscipt and marked it in red.}

\section{Editor in Chief}
Personally I like the paper, but it represents a somehow unusual
submission for Bayesian Analysis, and I am worried about it being of
sufficient general interest for the broad readership of the
journal. Still, in agreement with the AE, we decided to give the
author the possibility to revise the paper. However, I must stress
that there is no guarantee of a successful outcome, which will solely
depend on the quality of the revision. In particular:

\begin{description}
  \item[a] the Introduction is to be made accessible to the whole BA
    readership;
    \answer{I have considerably extended the introduction.}

  \item[b] the motivation and the practical statistical implications
    are to be developed and discussed;

    \answer{I added a section on the implications in the
      introduction.}

  \item[c] the contribution has to be better contextualized with
    respect to the literature in both inverse problems and D-optimal
    designs;
  \answer{adfasd}

  \item[d] numerical evidence and code, mentioned in the paper but
    actually not provided, need to be made available;

    \answer{I added a section describing numerical experiments and
    their results. I also added documentation to the accompanying
    \href{https://github.com/yairdaon/OED}{repository} and made sure
    all scripts there run.}

  \item[e] the mathematical derivations have to be watertight and some
    clarifications are needed in this respect (see Report of Referee
    3).

    \answer{I am not sure which specific comment of Reviewer 3 this
    refers to, please see my answers to Reviewer 3 below. }
\end{description}

\section{Associate Editor}
The three referees found merits and interest in your work, but also
pointed out several major issues to be addressed. Of particular
concern to me, the paper's exposition jumps too quickly into the
formulation / maths without giving a proper intuition and
context. This makes the paper less accessible to readers who aren't
already quite familiar with inverse problems (which are introduced
with very little preamble, and it's not even immediately explicit what
are the parameters being learned) and optimal experimental design.

\answer{I have added considerably to the introduction. See revised
manuscript and diff file.}

A referee makes valid points as to whether clusterization of
measurements is a problem in general, when in fact in can be an
appealing feature in some settings, and another referee asks whether
D-optimal designs should be abandoned altogether. The exposition
needs to discuss these high-level issues, and introduce the related
literature and results appropriately, to give the work proper
context. Altogether, I think that your work has the potential to make
a good contribution to Bayesian analysis, but as it currently stands
significant work is needed.

\section{Reviewer 1}
This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.

\answer{I appreciate the kind words.}

However, I feel changing the assumption from independent to correlated
measurements is quite artificial: people tweak a data generation
process assumption to solve an optimal design problem, and the author
just takes the assumption as it is.

\answer{If I understand correctly --- I completely agree. I added
text emphasizing that I do not endorse using correlated errors:}
\paste{asdf}

Since the author proves the clusterization issue under the independent
measurement assumption, I expect the author to go deeper and explain
why an ordinary data generation assumption + a common optimal design
leads to a “suboptimal” design?
\answer{Aagain, my apologies for the misunderstanding.}

Not just by the mathematical proof, but from an information theory
point of view. For example, does the redundancy points mean additional
points does not reduce the variance of the estimation at all? If
that’s the case, how does correlated measurement assumption “help”
with it?

\answer{Bla }

\section{Reviewer 2}
\subsection{Measurement Clusterization vs Repetition}
I will start by saying measurement clusterisation is not well-defined
in this paper. On page 1, it is referred to as two sets of
measurements that are almost indistinguishable from one
another. However, the setup for Corollary 11 implicitly defines
measurement clusterisation as repeated measurements. The latter
definition matches the references of Fedorov (1996) and Nyberg et al
(2012). So I will assume for the rest of this review that, measurement
clusterisation refers to repeated measurements.

\answer{I appreciate the comment, I fixed the relevant line in the
manuscript.}

Nearly all textbooks on classical design of experiments (e.g. Morris,
2011, Section 1.2.4) will feature a section with the three words
randomisation, blocking and repetition: the important principles of
design of experiments.  In these texts, repeated measurements are
considered to be a beneficial property of a design. Interestingly,
when optimal designs are found, they are often found to have repeated
measurements. The paper does not explain why the author thinks
repeated measurements (measurement clusterisation) is an undesirable
property.

\answer{I do not view clusterization as an undesired property, I only
mentioned many authors consider it to be undesirable. I have added a
section contrasting repetition and replication with clusterization and
why I believe they may be different. }

If one looks at, for example, Fedorov (1996) and Nyberg et
al (2012), it is an undesirable property because they are interested
in sensor location, and two or more sensors cannot be in the same
location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal

\answer{I believe that even in an application where repeated
experiments are allowed they may be unintuitive. For example, in a
medical imaging application --- taking a repeated slice in an MRI scan
might seem suboptimal to some. Of course, there is no argument
regarding the optimality give a model, which, as you mention might be
wrong.}

(if the model is correct, which it won’t be).

\answer{I believe clusterization may imply a model is somehow
  inadequate, see text below. }

An example, is a chemical engineering experiment
where one wishes to learn the relationship between a series of
variables and yield of a chemical reaction. The experiment will
involve specifying the variables, and measuring the yield (the
observation) after a specified period of time. This process is then
duplicated with a potentially different set of variables.  It is
perfectly possible to have repeated variables.

\answer{I couldn't agree more. Indeed, the suboptimality of clustered
design is a matter of personal opinion; once a design criterion and a
model are agreed upon there can be no disagreement.}

\begin{enumerate}

\item The author needs to make clear that they are considering
  experiments where repeated measurements are impossible. Admitedly,
  this consideration is implicit in the paper, for example, the
  examples given on page 1. However, it needs to be made explicit in
  the Abstract and the first paragraph of the paper.

  \answer{I do consider experiments where repeated measurements are
  possible. I am giving insight on why such designs are useful, even
  though repeated measurements may be non-intuitive in certain
  settings.}

\item The paper explains why repeated measurements can be optimal
  (answer to Question 3). The author should expand on this and draw
  the link with how repeated measurements are, typically, regarded as
  desirable in classical design of experiments for finite
  dimensions. See for example, the discussion of the relationship
  between Caratheodory’s Theorem and the number of support points in
  Pronzato \& Pazman (2013, page 139). The support points are the
  unique design points, so if this is less than m, then there are
  repeated design points.

  \answer{great comment}
\end{enumerate}

\subsection{Context}
The paper lacks discussion on context. I read the answers to the three
questions and thought “So what?”  What are the implications to
statistical practice, or experimental science, of these findings?  The
presentation of the paper does not help. The results are presented
with no explanation or intuition. It might be better to relegate much
of the mathematical detail to Supplementary Material and use the main
manuscript for exposition.

\answer{Thanks for this comment. I considerably expanded the
introduction and delegated many proofs to the Supplementary. I also
added a section on practical implications, reproduced below.}

\subsection{Avoiding measurement clusterisation}
The approaches to avoid measurement clusterisation, described on page
3, are described as pragmatic and "fundamentally altering the optimal
design problem".

Firstly, this is not strictly true. Suppose measurements correspond to
time, it is not possible to take two of more observations at the same
time, and there should be a minimum time period between measurements.
Implementing this minimum time period is not pragmatic, rather it is
characterising the application correctly.  One could imagine a similar
thing occurs with the eletrode locations on the skin in impedence
tomography: the physical size of the sensor imposes a miniumum
distance between sensors.

\answer{To a certain extend I agree: indeed, the impossibility of
taking identical measurements is indeed natural. However, this is not
encoded in the mathematical formulation of the problem. One can then
take one of two approaches: either accept what the mathematics
suggest, even if it counterintuitive, or find ways to circumvent the
mathematics because the result seems non-intuitive.}

Secondly, the paper seems to suggest “imposing correlations between
observations” as a solution. However, isn’t this imposition pragmatic
as well? I suppose it is argued that the $\obs \eps'$ term is
accounting for model error: but it is still pragmatic: the data
acquisition given by equation (1) is not the true data-generating
process.

\answer{I do not support imposing correlations in the paper, I just
used correlations to show how my framework can be used to rigorously
prove what is widely accepted. Other referees also understood that the
paper endorses imposing correlations and I accept that this is my
error in writing. I added a clause clearing the issue.}

\subsection{Minor comments}
\begin{itemize}
\item The author uses two different forms of asterisk: one for
  optimality of the design (e.g. Definition 3) and one for the
  adjoint. Suggest changing the one for optimality.

  \answer{I changed $\obs^\star$ to $\opt$.}

\item I do not like the use of the term measurements as used in the
  paper. The measurements are the $\data$’s since these are a result
  of measuring a physical quantity. However, the paper refers to the
  quantities that are controlled as the measurements.

  \answer{I was not able to find a better term to replace
  "measurements", so I stuck to it.}

\item Proof of Proposition 12 uses $\eps$ as part of the
  regularisation trick but this has been used previously for error
  (e.g. equation (1)).

  \answer{Absolutely correct, I changed it to $\nu$. You will not
  find the revised proof in the manuscript since I moved it to the
  Supplementray.}
\end{itemize}

\section{Reviewer 3}
The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior disribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clusterisation as the first main question being explored,
which is answered via, quoting, "randomized numerical simulations that
exhibit clusterization more than 95\% of the time (see the code in
supplementary material)". However, in the version of the supplementary
materials I had access to there is no further mention of these
numerical simulations and results; only a link to a Github repository
is provided, which is however not functioning.

\answer{I apologize for not properly maintainig the repository. It is
  now functioning and well-documented.}

Evidently, this issue needs to be fixed before any future
consideration can be given to the manuscript. For example, a detailed
description of the repeated experiments could be provided either in
the main article or in the supplement, with tables indicating the
various values of the parameters for the simulations and the precise
percentages obtained. Also, the code needs to be publicly available
for reproducibility.

\answer{I appreciate the suggestion. I added a section on numerical
simulations to the main atricle. It is reproduced below.}

\subsection{Further comments}
I have a number of further comments in regards to the results and the
overall presentation of the manuscript.

\begin{itemize}
\item D-optimal design vs. statistical recovery rates: being more
  familiar with the statistical inverse problem literature than with
  D-optimal design, the clusterisation phenomenon demonstrated in this
  paper raises the question as to wether consistent recovery of the
  unknown (m in the notation of this paper) is possible under this
  choice. All the results I am aware of either assume white
  noise/equally spaced design (e.g. [1]) or measurement loctions
  sampled uniformly at random (e.g. [2]). There are also some
  results with more general design under conditions on the
  fill-distance of the grid (e.g. [3]). All of these specifically
  prevent clusterisation, which is key to the statistical
  analysis. Hence, it is not clear to me even if D-optimal design
  should be pursued at all, if it indeed it leads to clusterisation as
  implied by the present paper. I think a discussion on this would be
  illuminating for the reader and help draw a connection to the
  broader statistical inverse problems literature.

\item Figure 1: It is hard to descern whether the measurements
  locations are perfectly overlapping or just very closely
  placed. Perhaps using dots with numbers here would help the
  readability.

  \answer{It is impossible to visually discern the measurements, even
  when resolution is increased. These measurements are not identical
  numerically, but they are identical to five decimal digits. Since I
  find these points via optimization over point location in
  $\Omega=[0,1]$ I view an error $<10^{-5}$ as reasonable.}

\item Experiment with the 1D heat equation: the experiment showcases
  some interesting phenomena, but also raises many questions. For
  instance: I would expect the dimensionality of the working domain
  (here the unit interval [0, 1]) to play a role, since larger
  dimension allow for higher freedom in placing the design points.

  \answer{That is a great point, but I view experiments in higher
    dimension as out of scope for the current study.}

\item Further, a dimensionality effect should also arise directly from
  Theorem 1 via the prior covariance eigenvalues: in the experiment
  the prior covariance operator is set to be equal to the inverse
  Laplace operator $(-\Delta)^{-1}$. In d-dimensional domains, its
  eigenvalues $\lambda_j$ are known to follow Weyl’s asymptotics,
  growing as $j^{2/d}$, which should then impact the index after which
  the eigenvalues are ‘thresholded’ by the procedure.

  \answer{Indeed, the growth of eigenvalues of the Laplacian is
    different in higher dimensions. This is why in higher dimensions
    we would have to take as prior \(u_0 \sim \mathcal{N}(\param_0,
    (-\Delta)^{-\gamma})\), for some \(\gamma > d/2\) in order to
    maintain regularity of prior realizations. I decided not to
    include a discussion on the choice of priors in higher spatial
    dimensions because it would require too much exposition and make
    me digress from the main points I wanted to convey, see the
    relevant paragraph below:}
  \paste{We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
    (-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
    choice ensures the posterior is well-defined and prior realizations
    are well-behaved, see \cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for
    details.}

\item Also, it should be relatively easy here to also study
  empirically the effect of the correlated error model in preventing
  clusterisation. Investigating all these issues in the context of the
  presented numerical set up would considerably enlarge the breath of
  the present work.

  \answer{I added relevant simulations to a new section Numerical
  Experiments.}

\item I found very hard to understand Theorem 1 immediately at the end
  of Section 1, as all the necessary background is introduced only
  later. Since the same result is also stated and proved in Section 5,
  I would consider cutting it from Section 1. The paragraphs in p.4
  and 5 already do a good job presenting the results, and the
  repetition of the formal statement does not seem necessary here.

  \answer{I removed the statement of Theorem 1 from the introudction.}

\item Denoting the optimal design in Theorem 1 by simply O is somewhat
  confusing, cf. the equation in the second item. Perhaps a separate
  notation such as Ō would help readability.

  \answer{Thank you for the comment, I completely agree. Optimal
  designs are now denoted $\opt$.}

\item Section 1.2 seems altogether unnecessary; the inverse regression
  model considered in the paper are a modelisation of many real-world
  phenomena, and they are routinely studied in statistical papers.

  \answer{I am glad you view this part as unnecessary, but I feel the
  model might feel too abstract to some practitioners. Thus, I choose
  to keep this part to avoid complaints from other readers.}

\item What does ‘strongly smoothning’ operator means in the thid line of Section 2.1?

  \answer{Basically I meant that its eigenvalues decrease
    "quickly". I understand this is not a common or well-defined term
    so I removed the relevant sentence.}

\item Section 4: the conclusion reached at the end of the section on
  p.15 seem to imply that in the noiseless limit, repeating any
  measurement does not improve the objective criterion, if correlated
  error model are present. Is it clear however that adding any
  (non-reapted) measurement always strictly increases it? I.e., there
  could be situations where adding one measurement does not change the
  value of the optimised criterion?

  \answer{Great comment! Such cases are possible but I view them as
  pathological. See short paragraph below.}

  Also, can the conclusion drawn here be extended to the
  case of noisy observations $\sigma > 0$?

  \answer{I do not expect this conclusion to hold: if noise is
  present, any measurement will result in some information gain and an
  increase in design criterion.}

\item At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
  defined for $\sigma^2= 0$, but in the latter case, could not a Gaussian
  posterior measure be still defined as in Gaussian process
  regression, e.g.[4].

  \answer{I believe a posterior can be defined, and so a KL
  divergence from said posterior to prior could be calculated and
  $\tar$ evaluated. However, taking $\sigma = 0$ in the main theorem
  implies $\tar \equiv \infty$. I believe this is because do not know
  how to mitigate this}

\item Theorem 1 is expressed in terms of the prior covariance matrix,
  which is a user specified quantity. Hence, it seems to me that all
  sort of design behaviour under the D-optimal criterion can occur by
  engineering the prior. It would perhaps be informative to study in
  more details some representative example of inverse problem and
  Gaussian prior. For example, the recovery of the initial condition
  with the ‘Matérn-like’ Gaussian prior laid out in the supplement are
  good candidates.

  \answer{the Matern-like prior actually arises by using a
  Laplacian-like operator as a prior, see Rue and Lindgren \cite{}. So
  a study of this example (with and without correlated errors) is
  already included in the paper. I am afraid implementing another
  inverse problem is out of scope for the present paper since I was
  not able to install common packages for inverse problems on my machine. }

  For these it is still not clear how the conclusion from Theorem 1
  translates into the Figure 1, also because I am unsure about the
  applicability of the result to point evaluations.

  \answer{Indeed, point evaluations $\delta_x$ are not in any
  function space I consider in the paper since my analysis is limited
  to Hilbert spaces. Point evaluations can be approximated well,
  though so in my opinion this is not a deal breaker. The only caveat
  is that upon approximating point evaluations e.g.~by $R(x;t) =
  t\mathbb{1}_{-t/2, t/2}(x)$ we will need to change the norm
  constraints to the norsomething different (i.e.~$\|R(\cdot;t\|$) but
  that does not change the analysis. I skim over this issue in the
  paper since I think it is confusing and adds little context.}

  In particular, how the conclusion that with $m = 4$ measurements the
  D-optimal procedure will aim to ignore the third and fourth
  eigenvalue/function was drawn?

  \answer{I explain this in more detail in the paper. See text below.}

\end{itemize}

\end{document}