### 2024-08-12 latex/implications.tex

Our answer to Question \ref{q:generic} implies that encountering
clusterization should be expected in many different problems across
many different scientific fields. Researchers that encounter
clusterization should not be surprised or wary. In Our answer to
Question \ref{q:why}, we explain what our view of the cause of
clusterization is. It appears, the cause is generic: a D-optimal
design reduces uncertainty for a select set of prior covariance
eigenvectors --- those with the most prior uncertainty, i.e.~those the
practitioner cares about the most! We believe practitioners should not
try to avoid measurement clusterization. Rather, practitioners should
take repeated measurements (e.g.~in MRI and borehole tomography),
increase apparatus sensitivity (e.g.~in EIT), or take consecutive
measurements (e.g.~in the 1D heat equation). Overall, we show that
measurement clusterization is a natural and (almost) inevitable part
of Bayesian D-optimal designs (but see disclaimer below).

One interesting implication of the analysis presented here is that
clusterization can serve as an evidence to the number of relevant
eigenvectors. Since leading eigenvectors typically correspond to slow
variations in space and/or time, clusterization could be used to
estimate the number of relevant degrees of freedom, and even to reduce
the complexity of a computational model, e.g.~by dropping
discretization points.

It is important to note that we do not view clustered designs as
undesirable, nor do we believe D-optimal designs should be avoided at
all. Nothing in our analysis indicates that we should expect any
pathological behavior when utilizing D-optimal designs. On the
contrary: we show that D-optimal designs (clustered or not) reduce
uncertainty of prior covariance eigenvectors where it is
highest. Thus, in a sense we hope to make precise in a future study,
we expect convergence of for D-optimal designs to be fastest.

For example, \cite{tekentrup2020} provides convergence analysis for
various designs with differing space filling properties. She showed
that convergence rate depends on how these designs fill space. We
expect better convergence rates for D-optimal designs; Even though
D-optimal desgins demonstrate clusterization, they do so because they
explore space in a way tailored to the problem studied. Consequently,
D-optimal designs cluster because they first aim to measure what
matters most; see the discussion following Theorem \ref{thm:char} for
details.

We expect D-optimal designs to judicious exploration of
design space, as implemented by a D-optimal design to give better
convergence compared to randomly sampling the design
space \cite{knapik2011}.

Clusterization is a peculiar phenomenon and it is perfectly reasonable
for someone to argue against the D-optimality criterion based on the
fact that it results in clustered designs. We have seen, however that
there is a perfectly reasonable explanation for clusterization. We
have shown that clusterization is an inevitable consequence of having
a problem with some modes where uncertainty decays faster than others.

Lastly, we believe that when clusterization arises, it should serve as
a warning sign to practitioners. In the inverse problem of the 1D heat
equation, clusterization occurs primarily because Laplacian
eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
should not occur: for small $T$, the heat distribution at $x_1$ should
give very little knowledge on the heat distribution at $x_2$. This
behavior stems from a well-known property of the heat equation: it
allows information to spread \emph{instantly} across the computational
domain \cite{renardy2006PDE}. In reality, heat (and information)
propagate at finite speeds. Of course, the known physical barrier for
information spread is the speed of light, but we expect heat to spread
considerably slower: heating an Olympic pool at one end should have no
immediate effect on the temperature at the other end.

Our choice of prior is also a potential major contributor to
clusterization. Our choice of Gaussian prior similarly implies
information is shared between distant locations in $\Omega$. Thus, we
suggest refraining from choosing Gaussian priors with inverse
Laplacian covariance operators. Rather, non-Gaussian priors could be
employed instead \cite{hosseini2017, hosseini2019}.

We believe that the emergence of clusterization in this context is
thus non-physical, arising from the way the inverse problems we
consider are phrased. Clusterization therefore indicates that the
underlying mathematical / Bayesian model is overly permissive and
fails to capture crucial physical constraints of the problem. We
suggest that when clusterization occurs, practitioners should consider
alternative models where information is localized in space and travels
at finite speed in the medium. Such models may not only provide more
physically accurate and meaningful results but may also mitigate the
issue of clusterization.