### 2024-08-12 latex/refs.tex

\documentclass{article}
\def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
\def\authors{Yair Daon}
\def\journal{Bayesian Analysis}
\def\doi{12345}

% Define title defaults if not defined by user
\providecommand{\lettertitle}{Author Response to Reviews of}
\providecommand{\papertitle}{Title}
\providecommand{\authors}{Authors}
\providecommand{\journal}{Journal}
\providecommand{\doi}{--}

\input{preamble}
\input{definitions}
\usepackage{xr}
\externaldocument{ba}

\begin{document}

% Make title
{\Large\bf \lettertitle}\\[1em]
{\huge \papertitle}\\[1em]
{\authors}\\
{\it \journal, }\texttt{doi:\doi}\\
\hrule

% Legend
\hfill {\bfseries RC:} \textbf{\textit{Reviewer Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text

I appreciate the many thoughtful comments by all of you. I believe
these comments helped me considerably improve the exposition of my
paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit. Please see below my answers to the
comments made by the referees, associate editor and editor in chief:
\rred{My answers are marked in red.} \bblue{In some places I copied
  text from the revised manuscipt and marked it in blue.}

\section{Editor in Chief}
\RC Personally I like the paper, but it represents a somehow unusual
submission for Bayesian Analysis, and I am worried about it being of
sufficient general interest for the broad readership of the
journal. Still, in agreement with the AE, we decided to give the
author the possibility to revise the paper. However, I must stress
that there is no guarantee of a successful outcome, which will solely
depend on the quality of the revision. In particular:

\RC the Introduction is to be made accessible to the whole BA
    readership;

\AR  I have considerably extended the introduction.

\RV the motivation and the practical statistical implications are to
be developed and discussed;

\AR I added a section on the implications:
\begin{quote}
  \input{implications}
\end{quote}

\RC the contribution has to be better contextualized with respect to
the literature in both inverse problems and D-optimal designs;

\RC numerical evidence and code, mentioned in the paper but actually
not provided, need to be made available;

\AR I added a section describing numerical experiments and their
results. I also added documentation to the accompanying
\href{https://github.com/yairdaon/OED}{repository} and made sure all
scripts there run.

\RC the mathematical derivations have to be watertight and some
clarifications are needed in this respect (see Report of Referee 3).

\AR I am not sure which specific comment of Reviewer 3 this
      refers to, please see my answers to Reviewer 3 below.

\section{Associate Editor}
\RC The three referees found merits and interest in your work, but also
pointed out several major issues to be addressed. Of particular
concern to me, the paper's exposition jumps too quickly into the
formulation / maths without giving a proper intuition and
context. This makes the paper less accessible to readers who aren't
already quite familiar with inverse problems (which are introduced
with very little preamble, and it's not even immediately explicit what
are the parameters being learned) and optimal experimental design.

\AR I have added considerably to the introduction. See revised
manuscript and diff file.

\RC A referee makes valid points as to whether clusterization of
measurements is a problem in general, when in fact in can be an
appealing feature in some settings, and another referee asks whether
D-optimal designs should be abandoned altogether. The exposition
needs to discuss these high-level issues, and introduce the related
literature and results appropriately, to give the work proper
context.

\RC Altogether, I think that your work has the potential to make a
good contribution to Bayesian analysis, but as it currently stands
significant work is needed.

\input{rev1}
\input{rev2}
\input{rev3}

\bibliographystyle{apalike}
\bibliography{../../lib.bib}

\end{document}