### 2024-08-22 latex/ba.tex
Only need to comment on tekentrup

% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
\input{definitions}

%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}

\startlocaldefs
% ** Local definitions **
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

%\title{\support{}}
\runtitle{Measurement Clusterization in D-optimal Designs}
%\author{Yair Daon}
%\numberwithin{equation}{section}

\begin{aug}
\author{\fnms{Yair} \snm{Daon}\thanksref{addr1}\ead[label=e1]{yair.daon@gmail.com}}
\runauthor{Yair Daon}

\address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
    \printead{e1} % print email address of "e1"
}

\end{aug}

\begin{abstract}
  Estimation of parameters in physical processes often demands costly
  measurements, prompting the pursuit of an optimal measurement
  strategy. Finding such strategy is termed the problem of
  \emph{optimal experimental design}, abbreviated as optimal
  design. Remarkably, optimal designs can yield tightly clustered
  measurement locations, leading researchers to fundamentally revise
  the design problem just to circumvent this issue. Some authors
  introduce error correlation among error terms that are initially
  independent, while others restrict measurement locations to a finite
  set of locations. While both approaches may prevent clusterization,
  they also fundamentally alter the optimal design problem.

  In this study, we consider Bayesian D-optimal designs, i.e.~designs
  that maximize the expected Kullback-Leibler divergence between
  posterior and prior. We propose an analytically tractable model for
  D-optimal designs over Hilbert spaces. In this framework, we make
  several key contributions: \textbf{(a)} We establish that
  measurement clusterization is a generic trait of D-optimal designs
  with independent Gaussian measurement errors, and prove that
  introducing correlations among measurement error terms mitigates
  clusterization. \textbf{(b)} We characterize D-optimal designs as
  reducing uncertainty across a subset of prior covariance
  eigenvectors. Finally, \textbf{(c)} We leverage this
  characterization to argue that measurement clusterization arises as
  a consequence of the pigeonhole principle: when more measurements
  are taken than there are locations where the select eigenvectors are
  large and others are small --- clusterization occurs.
\end{abstract}

%% ** Keywords **
\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd{35R30}
\kwd[; secondary ]{28C20}
\end{keyword}

\end{frontmatter}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{numerix}
%\input{discussion}

\begin{acks}[Acknowledgments]
  This study is a result of research I started during my PhD studies
  under the instruction of Prof.~Georg Stadler at the Courant
  Institute of Mathematical Sciences. I would like to thank him for
  his great mentorship, attention to details and kindness. I would
  also like to thank Christian Remling, who helped me find a proof for
  Lemma \ref{lemma:free} in
  \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
  Last, but certainly not least, I would like to thank the three
  referees, associate editor and editor in chief for providing
  detailed and insightful reviewes that have made this manuscript a
  whole lot better.
  %Having worked on this manuscript alone for several years, they have
  %almost become my coautoh

  This research was supported in part by an appointment with the
  National Science Foundation (NSF) Mathematical Sciences Graduate
  Internship (MSGI) Program sponsored by the NSF Division of
  Mathematical Sciences. This program is administered by the Oak Ridge
  Institute for Science and Education (ORISE) through an interagency
  agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
  is managed for DOE by ORAU. All opinions expressed in this paper are
  the author's and do not necessarily reflect the policies and views
  of NSF, ORAU/ORISE, or DOE.
  % This work was also supported by The Raymond and Beverly Sackler
  % Post-Doctoral Scholarship.
\end{acks}

\bibliographystyle{ba}
\bibliography{/Users/yairdaon/work//lib.bib}

\end{document}