### 2024-08-22 revision/refs.tex

\documentclass[draft]{article}
\def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
\def\authors{Yair Daon}
\def\journal{Bayesian Analysis}
\def\doi{12345}

% Define title defaults if not defined by user
\providecommand{\lettertitle}{Author Response to Reviews of}
\providecommand{\papertitle}{Title}
\providecommand{\authors}{Authors}
\providecommand{\journal}{Journal}
\providecommand{\doi}{--}

\input{preamble}
\input{../latex/definitions}
\usepackage{xr}
\externaldocument{../latex/ba}

\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}

\begin{document}

% Make title
{\Large\bf \lettertitle}\\[1em]
{\huge \papertitle}\\[1em]
{\authors}\\
{\it \journal, }\texttt{doi:\doi}\\
\hrule

% Legend
\hfill {\bfseries RC:} \textbf{\textit{Reviewer Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text

I appreciate the many thoughtful insightful comments by all of
you. These comments have helped me considerably improve the exposition
of my paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit. Please see below my answers to the
comments made by editor in chief (Section \ref{eic}), associate editor
(Section \ref{ae}) and three referees (Sections \ref{ref1}, \ref{ref2}
and \ref{ref3}).

%% \section{Editor in Chief}\label{eic}
%% \RC Personally I like the paper, but it represents a somehow unusual
%% submission for Bayesian Analysis, and I am worried about it being of
%% sufficient general interest for the broad readership of the
%% journal. Still, in agreement with the AE, we decided to give the
%% author the possibility to revise the paper. However, I must stress
%% that there is no guarantee of a successful outcome, which will solely
%% depend on the quality of the revision. In particular:

%% \RC the Introduction is to be made accessible to the whole BA
%%     readership;

%% \AR I have considerably extended the introduction.

%% \RC the motivation and the practical statistical implications are to
%% be developed and discussed;

%% \AR I added a section on the implications:

%% %\begin{quote}
%%   %\input{implications}
%% %\end{quote}

%% \RC the contribution has to be better contextualized with respect to
%% the literature in both inverse problems and D-optimal designs;

%% \RC numerical evidence and code, mentioned in the paper but actually
%% not provided, need to be made available;

%% \AR I added a section describing numerical experiments and their
%% results. I also added documentation to the accompanying
%% \href{https://github.com/yairdaon/OED}{repository} and made sure all
%% scripts there run.

%% \RC the mathematical derivations have to be watertight and some
%% clarifications are needed in this respect (see Report of Referee 3).

%% \AR I am not sure which specific comment of Reviewer 3 this refers to,
%% please see my answers to Reviewer 3 below.

%% \section{Associate Editor}\label{ae}
%% \RC The three referees found merits and interest in your work, but also
%% pointed out several major issues to be addressed. Of particular
%% concern to me, the paper's exposition jumps too quickly into the
%% formulation / maths without giving a proper intuition and
%% context. This makes the paper less accessible to readers who aren't
%% already quite familiar with inverse problems (which are introduced
%% with very little preamble, and it's not even immediately explicit what
%% are the parameters being learned) and optimal experimental design.

%% \AR I have added considerably to the introduction. See revised
%% manuscript and diff file.

%% \RC A referee makes valid points as to whether clusterization of
%% measurements is a problem in general, when in fact in can be an
%% appealing feature in some settings, and another referee asks whether
%% D-optimal designs should be abandoned altogether. The exposition
%% needs to discuss these high-level issues, and introduce the related
%% literature and results appropriately, to give the work proper
%% context.

%% \RC Altogether, I think that your work has the potential to make a
%% good contribution to Bayesian analysis, but as it currently stands
%% significant work is needed.

\input{rev1}
\input{rev2}
\input{rev3}

\bibliographystyle{apalike}
\bibliography{../../lib.bib}

\end{document}