### 2024-08-22 revision/rev1.tex

\section{Reviewer 1}\label{ref1}
\RC This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.

\AR I appreciate the kind words.

\RC However, I feel changing the assumption from independent to correlated
measurements is quite artificial: people tweak a data generation
process assumption to solve an optimal design problem, and the author
just takes the assumption as it is.

\AR If I understand correctly --- I completely agree. I added text
emphasizing that I do not endorse using correlated errors:

\begin{quote}
  Given the above discussion, it is our belief that the methods
  suggested by other authors to avoid clusterization are merely
  overlooking the problem. We do not view clustered designs as
  inherently bad on their own. Rather, we suggest that if a clustered
  design arises, then a practitioner should revisit their probabilistic
  and physical modeling assumptions. If the practitioner is confident in
  their assumptions, then clustered designs should be avoided only to
  the extent necessitated by the physical measuring apparatus.
\end{quote}

\RC Since the author proves the clusterization issue under the independent
measurement assumption, I expect the author to go deeper and explain
why an ordinary data generation assumption + a common optimal design
leads to a “suboptimal” design?

\AR In my view, this is a result of the unlocalized structure of
Laplacian eigenvectors. I believe this is more of a problem with our
physical models and choice of priors. See discussion below reproduced
from the revised manuscript:

\begin{quote}
  When clusterization arises, we believe it should serve as a warning
  signal to practitioners. In the inverse problem of the 1D heat
  equation, clusterization occurs primarily because Laplacian
  eigenvectors do not decay in $\Omega$. Consequently, measuring
  $u(x_1, T)$ at some point $x_1 \in \Omega$ provides information
  about $u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively,
  this should not happen: for small $T$, the heat distribution at
  $x_1$ should give very little knowledge on the heat distribution at
  $x_2$. This behavior stems from a well-known property of the heat
  equation: it allows information to spread instantly across the
  computational domain \cite{renardy2006PDE}. In reality, heat (and
  information) propagate at finite speeds. Of course, the known
  physical barrier for information spread is the speed of light, but
  we expect heat and information to spread considerably slower. Thus,
  clusterization may suggest that our physical model itself is at
  fault.

  Our choice of prior is perhaps an even greater contributor to
  clusterization. Our choice of Gaussian prior similarly implies
  information is shared between distant locations in $\Omega$. Thus,
  we suggest refraining from choosing Gaussian priors with inverse
  Laplacian covariance operators. Rather, non-Gaussian priors should
  be employed instead \cite{hosseini2017, hosseini2019}. So,
  clusterization may suggest we give more attention to the choice of
  prior.
\end{quote}

\RC Not just by the mathematical proof, but from an information theory
point of view. For example, does the redundancy points mean additional
points does not reduce the variance of the estimation at all? If
that’s the case, how does correlated measurement assumption “help”
with it?

\AR I think this is the same misunderstanding I commented on above;
For any nonzero observation error $\sigma^2 > 0$, repeated
measurements will indeed reduce the signal-to-noise ratio of the
measurement and result in an increase of the design criterion. I
briefly comment on this subject in the paragraph below:

\begin{quote}
  Clusterization should not be confused with replication. Replication
  requires that the experimentalist executes multiple trials under
  circumstances that are \emph{nominally identical} \cite[Section
    1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
  and even necessary aspect of optimal experimental design
  \cite{fisher1949design, morris2011,
    schafer2001replication}. Similarly to a design implementing
  replication, a clustered design reduces the signal-to-noise ratio of
  the repeated measurements \cite{telford2007brief}. The difference is
  that a clustered design takes repeated measurements at the expense
  of other quantities not measured at all.
\end{quote}