### 2024-08-25 revision/rev3.tex

\section{Reviewer 3}\label{ref3}
\RC The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior distribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

\RC While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clusterisation as the first main question being explored,
which is answered via, quoting, "randomized numerical simulations that
exhibit clusterization more than 95\% of the time (see the code in
supplementary material)". However, in the version of the supplementary
materials I had access to there is no further mention of these
numerical simulations and results; only a link to a Github repository
is provided, which is however not functioning.

\AR I made sure the repository is functioning and well-documented. See
\url{https://github.com/yairdaon/OED}.

\RC Evidently, this issue needs to be fixed before any future
consideration can be given to the manuscript. For example, a detailed
description of the repeated experiments could be provided either in
the main article or in the supplement, with tables indicating the
various values of the parameters for the simulations and the precise
percentages obtained. Also, the code needs to be publicly available
for reproducibility.

\AR I appreciate the suggestion. I added a section reproducing the
abovementioned numerical experiments to the main article. See Section
\ref{subsec:lemma_sims} reproduced below:

\begin{quote} %% Simulate
  In the proof of Theorem \ref{thm:char} we utilize Lemma
  \ref{lemma:free} to construct D-optimal designs. We implement this
  construction with the goal of testing numerically how prevalent are
  clustered designs. To this end, we would like to generate random
  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
  $\opt^*\opt$ and then utilize the construction of
  Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
  \newline
  To simplify things, we directly sample rank $\opt^*\opt$. We iterate
  over the number of measurements $m \in \{4,\dots, 24\}$, and for
  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
  m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
  times:
  \begin{enumerate}
  \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
    entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
    $\ttr D = m$.
  \item Conjugate $D$ by a random orthogonal matrix to form a positive
    semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
    represents $\opt^*\opt$.
  \item Apply the construction of Lemma \ref{lemma:free} to calculate
    $A$ such that $AA^t = M$, where $A$ has unit norm
    columns. I.e.~find the optimal design $\opt$.
  \item Since $A$ corresponds to $\opt$, its columns correspond to
    measurement vectors. We call $A$ "clustered" if $A$ has two or
    more identical columns (up to some numerical precision threshold).
  \end{enumerate}
  We then calculate the fraction of clustered designs of the
  simulations we ran, for each pair $m,k$. Clusterization occurred at
  high rates ($>99.9\%$) whenever $m-k > 1$; see
  Fig.~\ref{fig:sim_AAt}. Hence, in these simulations, clusterization
  is a generic property. However, when $m-k = 1$, clusterization does
  not occur. We do not why this is so and further investigation into
  this phenomenon is out of scope for the current study.
  \newline
  Full results are located in the \texttt{simulations.csv} file within
  the accompanying \href{https://github.com/yairdaon/OED}{repository}.
  Code implementing the experiments described above is located in
  module \texttt{zeros.py} of said repository. Runtime should be less
  than 30 minutes on any reasonably modern laptop (it took 12 minutes
  on the author's laptop).
\end{quote}

\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{../latex/figs/simulations.png}
    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
      generated randomly (see text and repository for details on
      generating $M$). It is evident that when $m-k \geq 2$ clusterization
      is ubiquitous, whereas for lower $m-k$ clusterization does not
      occur.}
  \label{fig:sim_AAt}
\end{figure}

\subsection{Further comments}
\RC I have a number of further comments in regards to the results and the
overall presentation of the manuscript.

\RC D-optimal design vs. statistical recovery rates: being more
familiar with the statistical inverse problem literature than with
D-optimal design, the clusterisation phenomenon demonstrated in this
paper raises the question as to whether consistent recovery of the
unknown ($\param$ in the notation of this paper) is possible under
this choice. All the results I am aware of either assume white
noise/equally spaced design (e.g.~\cite{knapik2011}) or measurement
locations sampled uniformly at random (e.g.~\cite{nickl2023}). There
are also some results with more general design under conditions on the
fill-distance of the grid (e.g.~\cite{teckentrup2020}). All of these
specifically prevent clusterisation, which is key to the statistical
analysis. Hence, it is not clear to me even if D-optimal design should
be pursued at all, if it indeed it leads to clusterisation as implied
by the present paper. I think a discussion on this would be
illuminating for the reader and help draw a connection to the broader
statistical inverse problems literature.

\AR I added a discussion on this subject. In the manuscript this
discussion is broken in two since the second part relies on Theorem
\ref{thm:char}. Thus, the first part is in the Introduction and the
second part follows the statement and proof of the abovementioned
theorem. I believe the level of discussion is adequate and view a
deeper analysis of convergence rates as out of scope for the current
study.
\begin{quote}%%Convergence
  Given that D-optimal designs suffer from clusterization, it is
  natural to ask whether D-optimal designs should be pursued at
  all. Other authors have established convergence results for decaying
  measurement error \cite{knapik2011}, space-filling measurements
  \cite{teckentrup2020} and randomly sampled designs
  \cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
  designs in particular --- do not fall into any of these categories.
  \newline
  Luckily, we can expect convergence from D-optimal designs (including
  clustered designs). Convergence holds in the following sense: the
  posterior uncertainty ellipsoid in $\hilo$ will contract to zero
  along every eigenfunction of $\fwd \prcov \fwd^*$. See Section
  \ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
  --- the main theorem of this manuscript.
  \newline
  Below we fulfill our promise from Section \ref{subsub:implications}
  and prove that the posterior uncertainty ellipsoid in $\hilo$ will
  contract to zero along every eigenfunction of $\fwd \prcov
  \fwd^*$. In our proof we ignore potential problems with conducting
  inference on function spaces which are not unique to D-optimal
  designs (see e.g.~\ref{owhadi2015} for more details).
  \newline
  First, let $\opt_m$ a D-optimal design utilizing $m$
  measurements. Then, an immediate consequence of Theorem
  \ref{thm:char} is that $\rank \opt_m^*\opt_m = k \to \infty$ as the
  number of measurements $m \to \infty$. Moreover, part (5) of Theorem
  \ref{thm:char} ensures that the eigenvalues of the posterior
  covariance equal
  \begin{equation*}
    \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^k \lambda_j^{-1} +
      \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$},
  \end{equation*}
  where we explicitly express the dependence of $k$ in the
  superscript. Now note that $\theta^{(k)}_i \leq \lambda_i$ for
  $i\leq k$. Specifically, for fixed $k$, we get $\theta^{(k)}_i \leq
  \lambda_k$ for all $i\leq k$. Since $\lim_{k\to \infty} \lambda_k=
  0$, we conclude that for all $i$, $\lim_{k\to\infty} \theta^{(k)}_i
  = 0$. Since $m\to \infty$ implies $k\to \infty$ we have
  $\lim_{m\to\infty} \theta^{(k)}_i = 0$ for all $i$. Therefore,
  posterior uncertainty decays to zero for all eigenvectors.
\end{quote}

\RC Figure 1: It is hard to discern whether the measurements locations
are perfectly overlapping or just very closely placed. Perhaps using
dots with numbers here would help the readability.

\AR It is impossible to visually discern the measurements, even when
resolution is increased. These measurements are not identical
numerically, but they are identical to five decimal digits. Since I
find these points via optimization over point location in
$\Omega=[0,1]$ I view a numerical error $<10^{-5}$ as reasonable.

\RC Experiment with the 1D heat equation: the experiment showcases
some interesting phenomena, but also raises many questions. For
instance: I would expect the dimensionality of the working domain
(here the unit interval $[0, 1]$) to play a role, since larger
dimension allow for higher freedom in placing the design points.
Further, a dimensionality effect should also arise directly from
Theorem 1\footnote{Now denoted Theorem \ref{thm:char}.} via the prior
covariance eigenvalues: in the experiment the prior covariance
operator is set to be equal to the inverse Laplace operator
$(-\Delta)^{-1}$. In d-dimensional domains, its eigenvalues
$\lambda_j$ are known to follow Weyl’s asymptotics, growing as
$j^{2/d}$, which should then impact the index after which the
eigenvalues are ‘thresholded’ by the procedure.

\AR These are two great points. Indeed, the growth of eigenvalues of
the Laplacian is different in higher dimensions. This is why in higher
dimensions we would have to take as prior \(u_0 \sim
\mathcal{N}(\param_0, (-\Delta)^{-\gamma})\), for some \(\gamma >
d/2\) in order to maintain regularity of prior realizations. I decided
not to include a discussion on the choice of priors in higher spatial
dimensions because (1) it would require too much exposition and make
me digress from the main points I wanted to convey and (2) due to
technical difficulties in installing packages for inverse problems
\cite{attia2023pyoed, VillaPetraGhattas16} I had to implement the
inverse problem of the 1D heat equation from scratch, so I decided to
leave out experiments in higher dimensions.

\RC Also, it should be relatively easy here to also study empirically
the effect of the correlated error model in preventing
clusterisation. Investigating all these issues in the context of the
presented numerical set up would considerably enlarge the breath of
the present work.

\AR I added relevant simulations to a new section Numerical
Experiments:

\begin{quote}
  In order to verify the results of Section
  \ref{section:non_vanishing}, we run simulations of the inverse
  problem of the 1D heat equation with nonvanishing model error
  \(\modcov = \prcov^2 \). Indeed, including model correlation pushes
  measurements apart, see Fig.~\ref{fig:corr_errors}. Code generating
  Fig.~\ref{fig:corr_errors} is located in module
  \texttt{clusterization.py} in the accompanying
  \href{https://github.com/yairdaon/OED}{repository}.
\end{quote}
\begin{figure}
  \centering
  \includegraphics[height=0.5\textwidth]{../latex/figs/dst_modelError4.png}
  \caption{Model correlation mitigates clusterization. We add a
    model correlation term to the error terms in the 1D heat
    equation inverse problem. Lo and behold, measurements are not
    close anymore and are pushed away thanks to the model error
    term.}
    \label{fig:corr_errors}
\end{figure}

\RC I found very hard to understand Theorem 1 immediately at the end
of Section 1, as all the necessary background is introduced only
later. Since the same result is also stated and proved in Section 5, I
would consider cutting it from Section 1. The paragraphs in p.4 and 5
already do a good job presenting the results, and the repetition of
the formal statement does not seem necessary here.

\AR I removed the statement of Theorem 1 from the introduction. It is
Theorem \ref{thm:char} now.

\RC Denoting the optimal design in Theorem 1 by simply $\obs$ is
somewhat confusing, cf.~the equation in the second item. Perhaps a
separate notation such as $\bar{\obs}$ would help readability.

\AR Thank you for the comment, I completely agree. Optimal designs are
now denoted $\opt$.

\RC Section 1.2 seems altogether unnecessary; the inverse regression
model considered in the paper are a modelisation of many real-world
phenomena, and they are routinely studied in statistical papers.

\AR I am glad you view this part as unnecessary, but I feel the model
might feel too abstract to some practitioners. Thus, I choose to keep
this part to avoid complaints from other readers.

\RC What does ‘strongly smoothing’ operator means in the third line of Section 2.1?

\AR Basically I meant that its eigenvalues decay "quickly". I
understand this is not a common or well-defined term so I removed the
relevant sentence.

\RC Section 4: the conclusion reached at the end of the section on
p.15 seem to imply that in the noiseless limit, repeating any
measurement does not improve the objective criterion, if correlated
error model are present.

\AR Indeed, this is very intuitive: when no observation error is
present, a repeated measurement does not tell us anything new.

\RC Is it clear however that adding any (non-repeated) measurement
always strictly increases it? I.e., there could be situations where
adding one measurement does not change the value of the optimised
criterion?

\AR Thanks for this insightful comment! Such cases are possible but I
view them as pathological. I added a short discussion reproduced
below:

\begin{quote}%% Pathologies
It is worth noting that by the nonnegativity of the KL divergence,
$\tar$ cannot decrease upon adding measurements. However, we can
construct examples where the posterior does not change upon taking a
new measurement e.g.~if the prior variance vanishes on some
eigenvector and a measurement is taken on said eigenvector. We do not
expect a measurement to generate no information gain whatsoever in any
realistic scenario, and ignore such pathologies.
\end{quote}

\RC Also, can the conclusion drawn here be extended to the case of
noisy observations $\sigma > 0$?

\AR I do not expect this conclusion to hold: if noise is present, even
a repeated measurement will result in some information gain and an
increase in design criterion. The reason is that for that particular
measurement the signal-to-noise ratio increases: observation error is
effectively reduced by a factor of $1/\sqrt{2}$ for the repeated
measurement. I discuss this briefly in the revised manuscript:

\begin{quote} %% Replication
Clusterization should not be confused with replication. Replication
requires that the experimentalist executes multiple trials under
circumstances that are \emph{nominally identical} \cite[Section
  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
and even necessary aspect of optimal experimental design
\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
to a design implementing replication, a clustered design reduces the
signal-to-noise ratio of the repeated measurements
\cite{telford2007brief}. The difference is that a clustered design
takes repeated measurements at the expense of other quantities not
measured at all.
\end{quote}

\RC At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
defined for $\sigma^2= 0$, but in the latter case, could not a
Gaussian posterior measure be still defined as in Gaussian process
regression, e.g~\cite{rasmussen2006}].

\AR Indeed, one can define a posterior, so a KL divergence from said
posterior to prior could be calculated and $\tar$ evaluated. However,
taking $\sigma = 0$ in Theorem 1 (the theorem of
\cite{AlexanderianGloorGhattas14}) gives $\tar \equiv \infty$. I did
not study this subject further.

\RC Theorem 1 is expressed in terms of the prior covariance matrix,
which is a user specified quantity. Hence, it seems to me that all
sort of design behavior under the D-optimal criterion can occur by
engineering the prior. It would perhaps be informative to study in
more details some representative example of inverse problem and
Gaussian prior. For example, the recovery of the initial condition
with the ‘Mat\'ern-like’ Gaussian prior laid out in the supplement are
good candidates.

\AR the Mat\'ern-like prior actually arises by using a Laplacian-like
operator as a prior \cite{rue2011}. So a study of this example (with
and without correlated errors) is already included in the paper. I am
afraid implementing another inverse problem is out of scope for the
present paper since I was not able to install common packages for
inverse problems \cite{attia2023pyoed, VillaPetraGhattas16} on my
machine.

\RC For these it is still not clear how the conclusion from Theorem 1
translates into the Figure 1, also because I am unsure about the
applicability of the result to point evaluations.

\AR Indeed, point evaluations $\delta_x$ are not in any function space
I consider in the paper since my analysis is limited to Hilbert
spaces. Point evaluations can be approximated well in
e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. For
example, we can replace point evaluations by $R(x;t) =
t\mathbb{1}_{-1/2t, 1/2t}(x)$, where $\mathbb{1}$ is an indicator
function. We will then need to change the norm constraints on
measurement vectors from unit to $\|R(\cdot;t\|^2 = \int R^2(x;t)dx =
t$ norm but that does not change the analysis as long as $t$ is
fixed. I barely mentioned this issue in the paper since I think it is
confusing and adds little context. See a footnote I added:

\begin{quote} %% Footnote
  The alert reader will likely ask how do we reconcile point
  measurements $\delta_x$ as suggested by the formulation of the 1D
  heat equation with working in Hilbert spaces. We don't. We follow
  standard practice in the literature and restrict our analysis to
  Hilbert spaces. We can satisfy ourselves with the fact that point
  evaluations could be approximated in a standard Hilbert space like
  $L^2(\Omega)$.
\end{quote}

\RC In particular, how the conclusion that with $m = 4$ measurements
the D-optimal procedure will aim to ignore the third and fourth
eigenvalue/function was drawn?

\AR I explain this in more detail in the revised manuscript:

\begin{quote}
  Building on Theorem \ref{thm:char}, we can now give a compelling
  explanation to the measurement clusterization we observed for the
  inverse problem of the heat equation, see Section
  \ref{subsub:clusterization1} below. We also suggest a generic
  explanation for clusterization, see Section
  \ref{subsub:clusterization2}.
  \newline
  Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
  heat equation}. As before, we denote the eigenvalues of
  $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
  our \emph{generic} model, and find a D-optimal design $\opt$ for our
  generic model using Theorem \ref{thm:char}. In our generic model,
  the measurements we take are best utilized in reducing uncertainty
  for the first $k$ eigenvectors. So, a D-optimal design arising from
  our \emph{generic model} completely avoids measuring eigenvectors
  $k+1$ and above.
  \newline
  Of course, in a real life problem --- such as the inverse problem of
  the 1D heat equation --- it is likely impossible to find measurement
  locations for which all eigenvectors $k+1$ and above are
  zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
  quickly (recall the square-exponential decay for eigenvalues of the
  1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
  try to balance measuring a small number (i.e.~$k$) of the leading
  eigenvectors.
  \newline
  The abovementioned balance is explored in
  Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
  = [0,1]$ and observe that D-optimal measurement locations are
  clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
  the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
  that eigenvectors $3$ and above have negligible prior
  amplitude. Since we only have $m=4$ measurements at our disposal, we
  interpret these results, following Theorem \ref{thm:char}, as
  implying we should only care about measuring the first and second
  eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
  compromise between the amplitude of the first and second
  eigenvectors. For example, a measurement at $x=0.5$ would have
  ignored the second eigenvector altogether, since the second
  eigenvector is zero at $x=0.5$.
  \newline
  Now we can understand measurement clusterization for the inverse
  problem of the heat equation. A D-optimal design attempts to measure
  the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
  (spatial) limitations on where these $k$ eigenvectors have large
  amplitude. For the inverse problem of the heat equation there are
  two spatial locations that present a good compromise between the
  amplitudes of the first and second eigenvectors, namely $x_1$ and
  $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
  measurements at our disposal but only two spatial locations that are
  a good compromise between the first and second scaled
  eigenvectors. Thus, clusterization arises as a consequence of the
  pigeonhole principle.
\end{quote}

\begin{figure}\label{fig:eigenvectors}
  \centering
  \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
  \caption{D-optimal measurement locations ($m=4$ measurements) and
    weighted eigenvectors for finding the initial condition of the 1D
    heat equation. Measurement locations and weighted eigenvectors are
    plotted over the computational domain $\Omega = [0, 1]$
    (x-axis). Measurement clusterization occurs approximately at
    $0.31$ and $0.69$. These two locations are a compromise between
    the magnitudes of the first and second eigenvectors, which are the
    eigenvectors that a D-optimal design aims to measure. Allocating
    $m=4$ measurements into two locations results in clusterization,
    according to the pigeonhole principle.}
  \label{fig:why}
\end{figure}