### 2024-08-25 revision/refs.tex
Now with refs

\documentclass{article}
\def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
\def\authors{Yair Daon}
\def\journal{Bayesian Analysis}
\def\doi{12345}

% Define title defaults if not defined by user
\providecommand{\lettertitle}{Author Response to Reviews of}
\providecommand{\papertitle}{Title}
\providecommand{\authors}{Authors}
\providecommand{\journal}{Journal}
\providecommand{\doi}{--}

\input{preamble}
\input{../latex/definitions}
\usepackage{xr}
\externaldocument{../latex/ba}

\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}

\begin{document}

% Make title
{\Large\bf \lettertitle}\\[1em]
{\huge \papertitle}\\[1em]
{\authors}\\
{\it \journal, }\texttt{doi:\doi}\\
\hrule

% Legend
\hfill {\bfseries RC:} \textbf{\textit{Referee Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text

I appreciate the many thoughtful insightful comments by all of
you. These comments have helped me considerably improve the exposition
of my paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit this manuscript. Please see below my
answers to the comments made by the editor in chief (Section
\ref{eic}), associate editor (Section \ref{ae}) and three referees
(Sections \ref{ref1}, \ref{ref2} and \ref{ref3}).

\section{Editor in Chief}\label{eic}
\RC Personally I like the paper, but it represents a somehow unusual
submission for Bayesian Analysis, and I am worried about it being of
sufficient general interest for the broad readership of the
journal. Still, in agreement with the AE, we decided to give the
author the possibility to revise the paper. However, I must stress
that there is no guarantee of a successful outcome, which will solely
depend on the quality of the revision. In particular:

\RC the Introduction is to be made accessible to the whole BA
    readership;

\AR I have considerably extended the introduction. See revised
manuscript, diff file and detailed answers to referees' comments
below.

\RC the motivation and the practical statistical implications are to
be developed and discussed;

\AR I added a section on the implications. The main takeaways are: (1)
avoid linearization and Gaussian priors, and (2) clusterization can
inform you of the number of relevant d.o.f in the problem.

\RC the contribution has to be better contextualized with respect to
the literature in both inverse problems and D-optimal designs;

\AR I added discussion on replication, contrasting it with
clusterization. I added convergence proof for D-optimal designs
(albeit without explicit rate) per Referee 2's request.

\RC numerical evidence and code, mentioned in the paper but actually
not provided, need to be made available;

\AR I added a section describing numerical experiments and their
results. I also added documentation to the accompanying
\href{https://github.com/yairdaon/OED}{repository} and made sure all
scripts there run.

\RC the mathematical derivations have to be watertight and some
clarifications are needed in this respect (see Report of Referee 3).

\AR I am not sure which specific comment of Referee 3 this refers to,
please see my answers to Referee 3 below.

\section{Associate Editor}\label{ae}
\RC The three referees found merits and interest in your work, but also
pointed out several major issues to be addressed. Of particular
concern to me, the paper's exposition jumps too quickly into the
formulation / maths without giving a proper intuition and
context. This makes the paper less accessible to readers who aren't
already quite familiar with inverse problems (which are introduced
with very little preamble, and it's not even immediately explicit what
are the parameters being learned) and optimal experimental design.

\AR I have added considerably to the introduction. I hope you will
find the transfer to the maths smoother now.

\RC A referee makes valid points as to whether clusterization of
measurements is a problem in general, when in fact in can be an
appealing feature in some settings,

\AR I added a discussion contrasting replication vs.~clusterization.

\RC and another referee asks whether D-optimal designs should be
abandoned altogether.

\AR I added comments on my suggestions to practitioners who encounter
clusterization: avoid Gaussian priors and linearization, separate
observations only to the extent required by the physical measurement
apparatus.

\RC The exposition needs to discuss these high-level
issues, and introduce the related literature and results
appropriately, to give the work proper context.

\AR I also added a discussion on convergence per Referee 2's request
contrasting D-optimal designs with other convergence results.

\RC Altogether, I think that your work has the potential to make a
good contribution to Bayesian analysis, but as it currently stands
significant work is needed.

\AR Again, I appreciate the time and effort invested by the referees
and editorial team. Thanks for the opportunity to revise and improve this manuscript.

\input{rev1}

\input{rev2}

\input{rev3}

\bibliographystyle{apalike}
\bibliography{/Users/yairdaon/work/lib.bib}

\end{document}